### 2024-10-22 manuscript.tex
Various improvements

Global storm-resolving models (GSRMs) are developed as the forthcoming phase of
climate modelling. Among them is a 5-km Icosahedral Nonhydrostatic Weather and
Climate Model (ICON). Due to the resolution, parametrisations of convection and
clouds are avoided. Standard-resolution models have substantial cloud biases
over the Southern Ocean (SO), affecting radiation and sea surface temperature.
We quantify how the GSRM ICON simulates clouds over the SO and contrast it with
the ECMWF Reanalysis 5 (ERA5) and Modern-Era Retrospective analysis for
Research and Applications, Version 2 (MERRA-2). The SO is dominated by boundary
layer clouds, which often cannot be observed properly by speceborne instruments
due to overlapping clouds, attenuation, and ground clutter. Instead, we use
about 2400 days of ship-based lidar observations from 31 voyages and one
subantarctic station. We use a ground-based lidar simulator for a like-for-like
comparison. ICON underestimates the total cloud fraction (TCF) by about 10\%,
with overestimation of clouds below 2 km and underestimation of clouds above 2
km. The reanalyses underestimate TCF by about 20\%. ERA5 overestimates clouds
below 1 km but underestimates near-surface clouds or fog. Compared to
radiosonde observations, ICON exhibits smaller natural variability, and its
lifting condensation level is higher, which might explain why cloud occurrence
peaks higher in the model than in observations. SO cloud biases are still a
substantial issue in a GSRM, even though an improvement over the
lower-resolution reanalyses is notable. The transition from parametrisation to
explicitly resolved convection and cloud processes might not be enough to
address the model cloud biases.

@@

We analysed a total of about 2400 days of lidar and radiosonde data from 31
voyages and campaigns and one subantarctic station data, covering the Atlantic,
Australian, and New Zealand sectors of the SO over the span of 10 years. This
dataset, together with the use of the ALCF lidar simulator, provided a fairly
comprehensive basis for evaluating SO cloud and thermodynamic profile biases in
the GSRM ICON and the ERA5 and MERRA-2 reanalyses. This provides a unique
evaluation perspective different from satellite observations, one that is more
suitable for evaluating bounary layer clouds, which are predominant in this
region. Furthermore, we subsetted our dataset by low and high latitude bands,
cyclonic activity, and stability in order to identify how these conditions
relate to the biases. Our main finding corroborates previous findings of large
boundary layer cloud biases in models and their subsequent effect on the
radiative transfer. This also applies to the new ICON GSRM, but the biases are
lower than in the reanalyses, despite the reanalyses having the advantage of
assimilation of the observed meteorological conditions.  The GSRM has, on the
other hand, the advantage of a much higher spatial resolution and, to a limited
extent, explicit calculation of traditionally subgrid-scale processes such as
convection.

The relationship between cloud biases and radiation has a number of notable
features. Perhaps unsurprisingly, the reanalyses exhibit the too few, too
bright bias previously identified in models. In our results, this is
characterised by outgoing TOA SW radiation similar to or higher than in
satellite observations, while at the same time total cloud fraction is
substantially underestimated relative to the ground-based lidar observations.
This feature seems to be much more pronounced in ERA5 than in MERRA-2. On the
other hand, this type of relationship is not present in ICON. This model mostly
predicts smaller outgoing TOA SW radiation and smaller total cloud fraction
than observations, and the deficit of outgoing TOA SW radiation is
approximately proportional to the deficit of the total cloud fraction. While this
might be a welcome feature and an improvement over previous models, it does
mean that the outgoing TOA SW radiation is overall underestimated instead of
being compensated by a higher cloud albedo. This can, of course, lead to
undesirable secondary effects such as overestimated solar heating of the sea
surface, among other factors responsible for SO SST biases in climate models
\citep{zhang2023,luo2023}.

We show that ICON underestimates the total cloud fraction by about 10\%, with
an overestimation of clouds below 2 km and an underestimation of clouds above
2 km. The reanalyses also underestimate the total cloud fraction by about 20\%.
ERA5 overestimates cloud below 1 km but underestimates near-surface cloud or
fog. In addition to lidar data, we compare radiosonde profiles acquired on the
RV \textit{Polarstern} voyages with ICON. Notably, the model exhibits smaller
natural variability than observations, and its lifting condensation level tends
to be higher. This might explain why cloud occurrence is peaking higher in the
model (at 500 m) than in observations (at the surface). Similar to our results,
\cite{cesana2022} showed that CMIP6 also tends to underestimate cloud
occurrence above 2 km over the SO, although their analysis in this case was
limited to liquid clouds.  We compared radiosonde profiles acquired on the RV
\emph{Polarstern} voyages with ICON.  The ICON model exhibits smaller internal
variability than observations, and its LCL level tends to be higher, which
might explain why a peak of cloud occurrence in observations is located at the
surface, while in ICON it is located higher, corresponding to the LCL.

@@

PK and FB received funding from the European Union’s Horizon 2020 research and
innovation program under a grant agreement no. 101003470, and supercomputing
resources were provided by the DKRZ (project 1125 ICON-development) and the
National Academic Infrastructure for Supercomputing in Sweden (allocation
2023/22-202).  We acknowledge the RV \emph{Polarstern} datasets provided by the
Alfred Wegener Institute and Pangaea, the AA15-16 dataset provided by the
Australian Antarctic Division (AAD) and University of Canterbury (UC), RV
\emph{Tangaroa} datasets provided by the National Institute of Water and
Atmospheric Research and UC, the NBP1704 dataset provided by the National
Science Foundation, Cooperative Institute for Research in Environmental
Sciences, University of Colorado and UC, the HMNZSW16 dataset provided by the
Royal New Zealand Navy and UC, the MARCUS dataset provided by the Atmospheric
Radiation Measurement (ARM) and AAD, and the MICRE dataset provided by ARM, the
Australian Bureau of Meteorology, and AAD. We thank the scientific staff, the
crew, and everyone involved in collecting data on the voyages and stations,
especially Kelly Schick, Peter Guest, John J.  Cassano, Simon Parsons, Graeme
Plank, Sally Garrett, Jamie Halla, Sean Hartery, Mike J.  Harvey (†), Simon P.
Alexander, and Alex Schuddeboom. We acknowledge the ICON model output provided
by the nextGEMS project and the DKRZ, reanalysis dataset ERA5 provided by the
Copernicus Climate Change Service and MERRA-2 provided by the Global Modeling
and Assimilation Office.  Last but not least, we acknowledge the use of open
source software Python, Cython \citep{behnel2011}, TensorFlow
\citep{abadi2016}, Devuan GNU+Linux, parallel \citep{tange2011}, NumPy
\citep{harris2020}, SciPy \citep{virtanen2020}, Matplotlib \citep{hunter2007},
cartopy \citep{cartopy}, pyproj, Inkscape, Bash, GNU Fortran, HDF
\citep{folk1999}, and NetCDF \citep{rew1990}.  We dedicate this study to the
memory of Mike J.  Harvey, who very substantially contributed to obtaining the
atmospheric observations on the RV \emph{Tangaroa} voyages used in this study.

>>>>

Global storm-resolving models (GSRMs) are the next avenue of climate modelling.
Among them is a 5-km Icosahedral Nonhydrostatic Weather and Climate Model
(ICON). Due to the resolution, parametrisations of convection and clouds are
avoided. Standard-resolution models have substantial cloud biases over the
Southern Ocean (SO), affecting radiation and sea surface temperature.  We
evaluated SO clouds in ICON and the ERA5 and MERRA-2 reanalyses.  The SO is
dominated by low clouds, which cannot be observed properly from space due to
overlapping clouds, attenuation, and ground clutter.  Instead, we analysed
about 2400 days of lidar observations from 31 voyages and a station using a
ground-based lidar simulator.  ICON and the reanalyses underestimate the total
cloud fraction by about 10 and 20\%, respectively. ICON and ERA5 overestimate
the cloud occurrence peak at about 500 m, potentially explained by their
lifting condensation level being too high.  The reanalyses strongly
underestimate near-surface clouds or fog.  MERRA-2 tends to underestimate cloud
occurrence at all heights. Unstable conditions are the most problematic for the
models. In daily cloud cover, ICON and the reanalyses tend to be about 1 and 2
oktas clearer, respectively. Compared to radiosondes, potential temperature is
accurate in all, but ICON is too unstable over low-latitude SO and too
moist near the surface. MERRA-2 is too humid at all heights. SO cloud biases
are a substantial issue in the GSRM, but it is an improvement over the
lower-resolution reanalyses. Explicitly resolved convection and cloud processes
were not enough to address the model cloud biases.

@@

We analysed a total of about 2400 days of lidar and 2300 radiosonde
observations from 31 voyages/campaigns and a subantarctic station, covering the
Atlantic, Australian, and New Zealand sectors of the SO over the span of 10
years. This dataset, together with the use of a ground-based lidar simulator,
provided a fairly comprehensive basis for evaluating SO cloud and thermodynamic
profile biases in the GSRM ICON and the ERA5 and MERRA-2 reanalyses. Our
analysis provides a unique evaluation perspective different from satellite
observations -- one that is more suitable for evaluating bounary layer clouds,
which are predominant in this region. Furthermore, we subsetted our dataset by
low and high latitude bands, cyclonic activity, and stability in order to
identify how these conditions relate to the biases.

Our main finding corroborates previous findings of large boundary layer cloud
biases in models and their subsequent effect on the radiative transfer. This
also applies to the new GSRM ICON, but the biases are lower than in the
reanalyses, despite the reanalyses having the advantage of assimilation of the
observed meteorological conditions.  The GSRM has, on the other hand, the
advantage of a much higher spatial resolution and, to a limited extent,
explicit calculation of traditionally subgrid-scale processes such as
convection.

We show that relative to ERA5, the distribution and strength of cyclonic
activity over the SO is well represented in ICON, but it is substantially less
stable in terms of LTS. The latter is also manifested in the radiosonde profile
comparison, which shows that the virtual potential temperature profiles in ICON
are more stable than in the observations over low-latitude SO.

The 31 voyages and a station show fairly similar biases in cloud occurrence by
height in the lidar comparison, which indicates that common underlying causes
for the biases exist regardless of longitude and season. ICON underestimates
the total cloud fraction by about 10\%, with an overestimation of clouds below
2 km and an underestimation of clouds above 2 km. The reanalyses also
underestimate the total cloud fraction by about 20\%.  ERA5 overestimates cloud
below 1 km but underestimates near-surface cloud or fog. ICON strongly
overestimates the peak of cloud occurrence at about 500 m, which might be
explained by the radiosonde comparison, which shows that it is too moist at
around this height.  Similar to our results, \cite{cesana2022} showed that
CMIP6 models also tend to underestimate cloud occurrence above 2 km over the
SO, although their analysis in this case was limited to liquid clouds.

Compared to the lidar observations, the daily cloud cover tends to be about 1
okta lower in ICON and 2 oktas lower in the reanalyses.  Unstable conditions
are associated with some of the greatest biases, especially in ERA5.  The
models also underestimate the cloud cover very strongly in cyclonic conditions,
which are very cloudy in the observations (8 oktas), but much less so in the
models.

The radiosonde observations indicate that the LCL is too high in ICON and
reanalyses, which is probably responsible for the higher peak of clouds in the
models and the lack of near-surface clouds or fog. The radiosonde comparison,
however, does not seem to explain cloud biases at higher altitudes.  MERRA-2 is
too moist at all heights.  ICON also exhibits smaller internal variability than
the radiosonde observations. Overall, the radiosonde comparison is only
partially explaining the identified cloud biases, and other physical causes are
likely contributing.  This warrants further investigation, especially of
ocean--atmosphere fluxes, shallow convection, and boundary layer turbulence.
The lack of parametrised subgrid-scale convection in ICON might be a
substantial issue even at the 5-km resolution.

The relationship between cloud biases and radiation has a number of notable
features. Perhaps unsurprisingly, the reanalyses exhibit the too few, too
bright bias previously identified in models. In our results, this is
characterised by outgoing TOA SW radiation similar to or higher than in the
satellite observations, while at the same time total cloud fraction is
substantially underestimated relative to the ground-based lidar observations.
This feature seems to be much more pronounced in ERA5 than in MERRA-2. On the
other hand, this type of relationship is not present in ICON. This model mostly
predicts smaller outgoing TOA SW radiation and smaller total cloud fraction
than observations, and the deficit of outgoing TOA SW radiation is
approximately proportional to the deficit of the total cloud fraction. While this
might be a welcome feature and an improvement over previous models, it does
mean that the outgoing TOA SW radiation is overall underestimated instead of
being compensated by a higher cloud albedo. This can, of course, lead to
undesirable secondary effects such as overestimated solar heating of the sea
surface, among other factors responsible for SO SST biases in climate models
\citep{zhang2023,luo2023}.

@@

PK and FB received funding from the European Union’s Horizon 2020 research and
innovation program under a grant agreement no. 101003470. Supercomputing
resources were provided by the DKRZ (project 1125 ICON-development) and the
National Academic Infrastructure for Supercomputing in Sweden (allocation
2023/22-202).  We acknowledge the RV \emph{Polarstern} datasets provided by the
Alfred Wegener Institute and Pangaea, the AA15-16 dataset provided by the
Australian Antarctic Division (AAD) and University of Canterbury (UC), RV
\emph{Tangaroa} datasets provided by the National Institute of Water and
Atmospheric Research and UC, the NBP1704 dataset provided by the National
Science Foundation, Cooperative Institute for Research in Environmental
Sciences, University of Colorado and UC, the HMNZSW16 dataset provided by the
Royal New Zealand Navy and UC, the MARCUS dataset provided by the Atmospheric
Radiation Measurement (ARM) and AAD, and the MICRE dataset provided by ARM, the
Australian Bureau of Meteorology, and AAD. We thank the scientific staff, the
crew, and everyone involved in collecting data on the voyages and stations,
especially Kelly Schick, Peter Guest, John J.  Cassano, Simon Parsons, Graeme
Plank, Sally Garrett, Jamie Halla, Sean Hartery, Mike J.  Harvey (†), Simon P.
Alexander, and Alex Schuddeboom. We acknowledge the ICON model output provided
by the nextGEMS project and the DKRZ, reanalysis dataset ERA5 provided by the
Copernicus Climate Change Service and MERRA-2 provided by the Global Modeling
and Assimilation Office.  Last but not least, we acknowledge the use of open
source software Python, Cython \citep{behnel2011}, TensorFlow
\citep{abadi2016}, Devuan GNU+Linux, parallel \citep{tange2011}, NumPy
\citep{harris2020}, SciPy \citep{virtanen2020}, Matplotlib \citep{hunter2007},
cartopy \citep{cartopy}, pyproj, Inkscape, Bash, GNU Fortran, HDF
\citep{folk1999}, and NetCDF \citep{rew1990}.  We dedicate this study to the
memory of Mike J.  Harvey, who very substantially contributed to obtaining the
atmospheric observations on the RV \emph{Tangaroa} voyages used in this study.
